---
title: "X Ray"
author: "Pablo Seibelt"
date: "22 de novembro de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### xray: The R Package to Have X Ray Vision on your Datasets

<https://blog.datascienceheroes.com/x-ray-vision-on-your-datasets/>
  
21 NOVEMBER 2017 on R, data analysis, machine learning, rstats, data cleaning
This package lets you analyze the variables of a dataset, to evaluate how is the shape of your data. 

Consider this the first step when you have your data for modeling, you can use this package to analyze all variables and check if there is anything weird worth transforming or even avoiding the variable altogether.

####Installation

You can install xray from github with:

```{r}
# library(httr)
# set_config(config(ssl_verifypeer = 0L))
# 
# devtools::install_github("sicarul/xray")
```


---

  
####Usage

#####Anomaly detection

xray::anomalies analyzes all your columns for anomalies, whether they are NAs, 
Zeroes, Infinite, etc, and warns you if it detects variables with at least 80% 
of rows with those anomalies. It also warns you when all rows have the same value.

Example:

 
```{r}
data(longley)  
badLongley=longley  
badLongley$GNP=NA  
xray::anomalies(badLongley) 
```

---

#####Distributions

xray::distributions tries to analyze the distribution of your variables, 
so you can understand how each variable is statistically structured. 

It also returns a percentiles table of numeric variables as a result, which can 
inform you of the shape of the data.

```{r}
distrLongley=longley  
distrLongley$testCategorical=c(rep('One',7), rep('Two', 9))  
xray::distributions(distrLongley)  
```

---

#####Distributions along a time axis

xray::timebased also investigates into your distributions, but shows you the 
change over time, so if there is any change in the distribution over time 
(For example a variable stops or starts being collected) you can easily visualize it.

```{r}
dateLongley=longley  
dateLongley$Year=as.Date(paste0(dateLongley$Year,'-01-01'))  
dateLongley$Data='Original'  
ndateLongley=dateLongley  
ndateLongley$GNP=dateLongley$GNP+10  
ndateLongley$Data='Offseted'  
xray::timebased(rbind(dateLongley, ndateLongley), 'Year') 
```




If you are just starting to learn about Data Science or want to explore further, 
i encourage you to check this cool book made by my buddy Pablo Casas: 

The Data Science Live Book

<https://livebook.datascienceheroes.com/>

---
